Errata

### 2025-04-05 tex/Manuscript-3.tex
Remove auxiliary files for Manuscript-4, Manuscript-5, Methodology, and Results_Discussion to clean up the repository and prevent unnecessary clutter.

>>>>

\begin{center}
    \textbf{\Large Manuscript 2}

    \vspace{0.3cm}
    \textit{Frontiers in Microbiology}\\
    \textit{DOI 10.3389/fmicb.2024.1319997}

    \vspace{0.5cm}

    Nicolás Riveras-Muñoz\textsuperscript{1*}, Steffen Seitz\textsuperscript{1}, Corinna Gall\textsuperscript{1}, Victoria Rodríguez\textsuperscript{2}, Kristina Witzgall\textsuperscript{3}, Peter Kühn\textsuperscript{1}, Carsten W. Mueller\textsuperscript{4,5}, Rómulo Oses\textsuperscript{6}, Oscar Seguel\textsuperscript{7}, Dirk Wagner\textsuperscript{2,8}, and Thomas Scholten\textsuperscript{1}

    Victoria Rodríguez\textsuperscript{1 *}, Alexander Bartholomäus\textsuperscript{1}, Kristina Witzgall\textsuperscript{2}, Nicolás Riveras-Muñoz\textsuperscript{3}, Romulo Oses\textsuperscript{4}, Susanne Liebner\textsuperscript{1,5}, Jens Kallmeyer\textsuperscript{1}, Oliver Rach\textsuperscript{6}, Carsten W. Mueller\textsuperscript{7,8}, Oscar Seguel\textsuperscript{9}, Thomas Scholten\textsuperscript{3} and Dirk Wagner\textsuperscript{1,10}

    \vspace{0.2cm}
  \end{center}

  \begin{small}
    \begin{justify}

        \textsuperscript{1}GFZ German Research Centre for Geosciences, Section Geomicrobiology, Potsdam, Germany,

        \textsuperscript{2}Soil Science, TUM School of Life Sciences, Technical University of Munich, Freising-Weihenstephan, Germany,

        \textsuperscript{3}Department of Geosciences, Soil Science and Geomorphology, University of Tübingen, Tübingen, Germany

        \textsuperscript{4}Centro Regional de Investigación y Desarrollo Sustentable de Atacama (CRIDESAT), Universidad de Atacama, Copiapó, Chile

        \textsuperscript{5}Institute of Biochemistry and Biology, University of Potsdam, Potsdam, Germany

        \textsuperscript{6}GFZ German Research Centre for Geosciences, Section Geomorphology, Potsdam, Germany

        \textsuperscript{7}Institute for Ecology, Chair of Soil Science, Technische Universitaet Berlin, Berlin, Germany

        \textsuperscript{8}Department of Geosciences and Natural Resource Management, University of Copenhagen, Copenhagen, Denmark

        \textsuperscript{9}Facultad de Ciencias Agronómicas, Universidad de Chile, Santiago, Chile

        \textsuperscript{10}Institute of Geosciences, University of Potsdam, Potsdam, Germany

    \end{justify}
  \end{small}

  \vspace{0.5cm}
  \begin{center}
    Received: 11 October 2023\\
    Accepted: 02 January 2024\\
    Published: 17 January 2024
  \end{center}
  \cleardoublepage

\section*{Abstract} % Use \section* for unnumbered sections if needed

The microbiota is attributed to be important for initial soil formation under extreme climate conditions, but experimental evidence for its relevance is scarce. To fill this gap, we investigated the impact of in situ microbial communities and their interrelationship with biocrust and plants compared to abiotic controls on soil formation in initial arid and semiarid soils. Additionally, we assessed the response of bacterial communities to climate change. Topsoil and subsoil samples from arid and semiarid sites in the Chilean Coastal Cordillera were incubated for 16 weeks under diurnal temperature and moisture variations to simulate humid climate conditions as part of a climate change scenario. Our findings indicate that microorganism-plant interaction intensified aggregate formation and stabilized soil structure, facilitating initial soil formation Interestingly, microorganisms alone or in conjunction with biocrust showed no discernible patterns compared to abiotic controls, potentially due to water masking effects. Arid soils displayed reduced bacterial diversity and developed a new community structure dominated by Proteobacteria, Actinobacteriota, and Planctomycetota, while semiarid soils maintained a consistently dominant community of Acidobacteriota and Proteobacteria. This highlighted a sensitive and specialized bacterial community in arid soils, while semiarid soils exhibited a more complex and stable community. We conclude that microorganism-plant interaction has measurable impacts on initial soil formation in arid and semiarid regions on short time scales under climate change. Additionally, we propose that soil and climate legacies are decisive for the present soil microbial community structure and interactions, future soil development, and microbial responses

\section*{Keywords} % Use \section* for unnumbered sections if needed

arid soil, semiarid soil, manipulation experiment, climate change, initial soil formation,
bacterial community.

\section{Introduction}
\label{sec:introduction}

Soil formation occurs at the dynamic interface between the atmosphere, hydrosphere, and lithosphere, playing a vital role in supporting life on Earth and facilitating essential biogeochemical cycles \citep{Amundson2007}. It involves interactions among physical, chemical, and biological processes, encompassing the weathering of parent rock materials and incorporating organic matter produced by biota \citep{Schulz2013}. In the early stages of soil formation, well-adapted microorganisms colonize bare mineral substrates, establishing simple communities that undergo successional changes toward more complex species interactions closely linked to soil formation \citep{Bajerski2013, Schulz2013, Liu2019, Li2021, Yu2022}. Concurrently, the presence of lichens and bryophytes, along with the initial assembly of microbial communities (bacteria, fungi, and algae), leads to the formation of biological soil crusts (BSCs). These BSCs influence the nutrient input through photosynthesis, nitrogen (N) fixation, and dust capture \citep{Elbert2012} while stabilizing the soil structure and protecting against erosion \citep{Seitz2017, RiverasMunoz2022}. As soil complexity and nutrient availability increase, vascular plants establish and promote weathering processes \citep{Calvaruso2009} and aggregate formation, which stabilizes soil structure jointly with roots in the rhizosphere and organic carbon accumulation \citep{Erktan2016, Baumert2018}. Understanding the initial role, assembly, and succession of microbial communities and their intimate associations with BSCs and vascular plants is crucial for comprehending the development of soils.

Research on the role and succession dynamics of microbial communities in arid soils with low nutrient content is limited compared to low biomass soils at volcanic eruption sites or glacier forefields, which have been more extensively studied \citep{Bajerski2013, Kelly2014, Meier2019, Hernandez2020}. Several studies have provided compelling evidence of the role of specific microbial species in soil development through carbon and N-fixation \citep{Schmidt2008, Duc2009}, efficient mineral weathering \citep{Frey2010, Lapanje2012}, and aggregate stabilization \citep{Kohler2006}. Furthermore, the presence of specific taxa and their predicted functions have been linked to these processes during soil development \citep{Zhelezova2019, Krauze2021, Rodriguez2022}, suggesting the significance of microbial communities in soil formation. Nevertheless, experimental evidence on the interplay between the natural microbial community and soil formation processes remains limited. For instance, investigations have shown that microbial inoculum from arable topsoil can enhance soil aggregation in artificial soils facilitated by exopolysaccharides and fungal hyphae \citep{Pronk2012}. Additionally, ubiquitous microorganisms of mediterranean soils with moisture reaching 10\% of the water-holding capacity have resulted in a tenfold increase in aggregation compared to sterile conditions \citep{Blankinship2016}. However, there is a significant gap in experimental evidence regarding the impact of microbial communities in succession from initial bare soils under arid and semiarid climate conditions into soils with a plant-root system.

Such arid and semiarid environments, which cover approximately 30\% of the Earth’s land surface, typically experience slow soil development due to limited soil moisture and water movement for material dissolution and leaching \citep{Abdelfattah2013, Sousa2021}. However, changes in precipitation patterns can have complex effects on vegetation, chemistry, mineralogy, and soil stability \citep{Su2010, Wang2017, Sousa2021}. Consequently, although soil formation is a long-term process over hundreds and thousands of years, it can be largely accelerated under optimal water availability and temperature conditions \citep{Buchan2011, Szalai2021}. Moreover, climate projections for the upcoming decades indicate increased extreme precipitation events in arid regions \citep{Donat2016, Ortega2019}. Therefore, there is a growing interest in investigating how modifications in temperature and precipitation influence soil formation processes in arid soils for predicting the impacts of climate change or improving land management practices \citep{Gelybo2018}.

Moisture and temperature changes notably impact soil microorganisms, particularly in arid and semiarid environments. Increased or decreased water availability affects gene copy numbers \citep{Barnard2013}, respiration and productivity \citep{Armstrong2016}, biomass \citep{Na2019, Ouyang2020}, diversity, and community composition \citep{Maestre2015, StovicekB2017}, due to alterations in soil properties (such as soil organic carbon content) and plant biomass. Key enzymatic activities in the carbon cycle, such as cellobiohydrolase and \textbeta-glucosidase \citep{Ouyang2020}, and in the phosphorous metabolism, such as acid phosphatase \citep{Na2019} also respond to water availability. However, the microbial response depends on the intensity and frequency of water pulses and ambient temperature, highlighting the complexity of bacterial responses \citep{StovicekA2017}. For instance, sudden and substantial water input in hyperarid regions can harm many soil microbial species and decrease diversity \citep{AzuaBustos2018}, while in other studies in the desert, it did not affect diversity \citep{Armstrong2016}. As our understanding of the direction and magnitude of microbial feedback to climate change continues to advance, it becomes increasingly important to investigate their implications for soil formation.

The Chilean Coastal Cordillera offers a unique opportunity to study soil formation across a climate gradient ranging from arid to humid conditions. Specifically, the arid site has drawn significant interest as a natural laboratory due to its extreme climate conditions and unique microbial community adapted to the harsh environment \citep{Genderjahn2018, SchulzeMakuch2018, Aguilera2021, Genderjahn2021, SchulzeMakuch2021}. Moreover, the arid and semiarid sites of the Chilean Coastal Cordillera share similarities in tectonic, lithologic, and pedologic factors, as well as low nutrient content and precipitation levels \citep{Bernhard2018, Oeser2018}. This makes them suitable for investigating the effects of humid conditions on initial soil formation and microorganisms.

This study tests the hypothesis that soil microbial communities, besides abiotic factors, plants, and biological soil crusts, directly speed up short-term soil formation in arid regions, especially when humidity and temperature are favorable. Additionally, we propose that soil microbial communities exhibit distinct responses to simulated climate change, with variations influenced by the soil legacy. We conduct an experimental study simulating humid conditions of the southern Coastal Cordillera and implement four treatments representing a successional gradient of soil development: sterile soil, soil with in situ microorganisms, soil with in situ microorganisms and BSCs, and soil with in situ microorganisms and plants. The impact of microbiota, BSCs, and plants on soil formation dynamics over time was evaluated by quantifying changes in soil physicochemical properties in the different treatments. We also characterize the diversity, structure, and interactions of the bacterial community, providing insights into their response to environmental changes.

\section{Materials and methods}
\subsection{Study sites and sample collection}

We collected soil samples from two protected areas, Pan de Az{\'u}car National Park (PA) and Santa Gracia Natural Reserve (SG), located in the Chilean Coastal Cordillera (\cref{fig:figure_label}, Supplementary \cref{fig:supp_figure_label}). 
The lithology at both sites is granitoid, ensuring similar parent material for soil formation and a meaningful comparison of the obtained results \citep{Bernhard2018, Oeser2018}. 
The mean annual precipitation and mean annual temperature are \SI{12}{\milli\metre} and \SI{16.8}{\degreeCelsius} in PA and \SI{66}{\milli\metre} and \SI{13.7}{\degreeCelsius} in SG, respectively \citep{Munoz2007}. 
These climatic differences result in distinct soil moisture and temperature regimes. 
PA is arid with almost no precipitation and mostly endorheic water resources. 
In contrast, SG is in a semiarid zone with coastal fog influence, annual rainfall variation, and winter rain \citep{Bernhard2018}.

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M3-Figure_1.jpg}
	\caption{Sampling sites and microcosms of Pan de Azúcar and Santa Gracia after 12 weeks of incubation. The different microcosms are separated by horizon and treatment, including in situ, BSC (Biocrust), and plant treatment.}
	\label{fig:M3-F1}
\end{figure}

The field sampling campaign took place in March 2019, at the end of the austral summer season. The coordinates and elevations for PA were \ang{26;06;35}S, \ang{70;32;57}W, and \(\sim\)\SI{336}{\meter} elevation, while for SG, they were \ang{29;45;15}S, \ang{71;10;3}W, and \(\sim\)\SI{720}{\meter} elevation. Soil samples were taken from five plots devoid of vegetation and two soil horizons (A and B) for each site (Supplementary \cref{tab:supp_table_label}). In PA, the A horizon (PA - Ahor) ranged from \SIrange{0}{2}{\centi\metre}, and the B horizon (PA - Bhor) ranged from \SIrange{2}{25}{\centi\metre}. Similarly, in SG, the A horizon (SG - Ahor) ranged from \SIrange{0}{3}{\centi\metre}, and the B horizon (SG - Bhor) ranged from \SIrange{3}{40}{\centi\metre}. The soil structure characteristics varied between horizons. PA - Ahor exhibited a cemented crust, while PA - Bhor had a few round aggregates between coarse material, single grain dominating in the fine fraction, and gypsum concentrated in one of the plots. SG - Ahor was characterized by weak aggregates, and SG - Bhor ranged from weak to strong and stable aggregates. Fresh samples from each plot and horizon separately were homogenized through a \SI{2}{\milli\metre} sieve and stored at \SI{4}{\degreeCelsius} until use.

\subsection{Soil manipulation experiment}

An experiment was performed to stimulate initial soil formation into developed soil ecosystems by applying humid climate conditions in soil microcosms, mimicking a climate change scenario. The simulated climate change was based on the climate data of Nahuelbuta National Park in the Chilean Coastal Cordillera, where we can find well-developed deep soils with a stable soil structure \citep{Bernhard2018, RiverasMunoz2022}. As in the natural conditions of Nahuelbuta, we applied a photoperiod of \SI{14}{\hour}/\SI{10}{\hour} day/night and the daily temperature fluctuation of January (\SI{18.3}{\degreeCelsius}/\SI{11.2}{\degreeCelsius} day/night) in the soil manipulation experiment. To simulate water availability by precipitation, \SI{25}{\milli\liter} of filtered and autoclaved MiliQ water was gently applied at once three days per week until 65\% water-filled pore space was reached after each application. This value approximates the precipitation in Nahuelbuta (\SI{1469}{\milli\metre\,\year^{-1}}) but is lower due to the water retention capacity of desert soils. The microcosms consisted of \SI{130}{\gram} of field moist soil (sieved $<$\SI{2}{\milli\metre} and previously stored at \SI{4}{\degreeCelsius} for nine months) placed in sterile polyvinyl chloride columns (\SI{70}{\milli\metre} in height and \SI{50}{\milli\metre} in diameter) with a silk fabric underneath to prevent soil scattering. The microcosms were arranged on a grid inside an incubator (Memmert ICP-110, Germany), mimicking the daily temperature fluctuations. The photoperiod was set using LED-SANlight S2W (IMAGRO, Germany) and controlled with a timer coordinated with temperature fluctuations.

For the manipulation experiment setup, soil material from five plots representing top and subsoil was pooled and homogenized. The experiment included 144 microcosms representing two sites (PA and SG), two soil horizons (Ahor and Bhor), four treatments, three sampling times, and three biological replicates for each combination (Supplementary \cref{fig:supp_figure_s2_label}). In addition, three original, untreated samples were collected as baseline measurements for each site and horizon (T0), totaling 156 samples. The treatments consisted of a successional gradient of soil development as follows: (i) in situ samples with indigenous microorganisms but without BSCs or plants, (ii) in situ samples with naturally growing BSCs to assess the impact of microorganisms at a later stage of soil development, (iii) in situ samples with \textit{Helenium aromaticum} (Hook.) L.H. Bailey (pioneer plant, described below) to simulate later stages of soil development and (iv) abiotic controls based on soil sterilized samples with \ce{^{60}Co} as a source of gamma rays and a dose of \SI{50}{\kilo Gy} (BGS Beta–Gamma–Service GmbH and Co. KG, Wiehl, Germany) to exclude the impact of indigenous microorganisms. \textit{H. aromaticum} is a native vascular annual herb widely distributed in Chile, known as a pioneer plant that colonizes disturbed soils, forming seasonal grasslands \citep{GomezGonzalez2011}. For plant preparation, the seeds were sterilized using 20\% v/v sodium hypochlorite and 0.05\% Tween 20, washed three times with autoclaved MiliQ water, and placed in sterile petri dishes with absorbent paper for germination. The germinated seeds were subsequently transferred to the microcosms. BSCs occurred naturally in the soil, whereas in situ samples with microorganisms were covered with a black breathable cloth to inhibit their growth.

Microcosms were sampled at 2 weeks (T1), 12 weeks (T2), and 16 weeks (T3) of incubation. Forty eight samples were sacrificed for each sampling time, covering two sites, two soil horizons, four treatments, and three replicates. In the 12th week of sampling, the plants were harvested, and the soil with the roots was retained until the 16th week of the experiment to assess their effects on the soil. Three microcosms were transferred to a laminar flow chamber for each treatment and time point to ensure sterile conditions during sub-sampling. From the total sample, \SI{80}{\gram} of soil was collected for soil physicochemical analysis and stored in the dark at room temperature. Additionally, \SI{10}{\gram} of soil was taken for DNA extraction and held at \SI{-20}{\degreeCelsius}, while \SI{40}{\gram} was stored at \SI{-80}{\degreeCelsius} for future RNA extraction and subsequent analysis.

\subsection{Soil physicochemical properties}

To determine initial soil formation, the aggregate size distribution, and the content of total carbon (C\textsubscript{t}) and total nitrogen (N\textsubscript{t}) by aggregate size, we collected a total of 104 soil samples from three different sampling points (T0, T1, and T3). For each sample, \SI{5}{\gram} of dry soil was shaken using a modified Casagrande apparatus (Mennerich Geotechnik, Hannover, Germany). The apparatus consisted of two sieves with mesh sizes of \SI{250}{\micro\metre} and \SI{53}{\micro\metre}, placed on top of a collecting vessel. The soil samples were subjected to 1,000 repeated cycles of sieving at a frequency of \SI{2}{\hertz}, resulting in the separation of three aggregate size classes: macroaggregates (\SI{> 250}{\micro\metre}), large microaggregates (\SIrange{53}{250}{\micro\metre}), and small microaggregates and primary particles (\SI{< 53}{\micro\metre}). The data on aggregate size distribution was used to estimate aggregate stability given by the mean weight diameter of the aggregates ($\Delta$MWD) as follows:

    $$\Delta MWD = \frac{\sum_{i=1}^{n} X_i*W_i}{\sum_{i=1}^{n} W_i}$$

\(W_i\) is the corrected mass proportion of stable aggregate fraction \(i\) in the total aggregates (\SI{< 2}{\milli\metre}) and \(X_i\) is the mean diameter of stable aggregate fraction \(i\).

Subsequently, ball milling was conducted before measuring the content of C\textsubscript{t} and N\textsubscript{t} in all aggregate size classes. The process involved milling the samples for \SI{3}{\minute} at an oscillation frequency of \SI{30}{\hertz}. C\textsubscript{t} and N\textsubscript{t} measurements for all fraction samples were performed using a CHNSO elemental analyzer (HEKAtech Euro EA, Wegberg, Germany).

Additionally, we determined the total organic carbon (TOC) and total nitrogen in the bulk soil (N\textsubscript{bulk}) without aggregate size separation. This analysis used \SI{6}{\milli\gram} of bulk soil from the same subset of soil samples (104). The measurements were carried out using a flash HT 2000 Organic elemental analyzer, following a \SI{6}{M} chlorhydric acid digestion of carbonate at \SI{70}{\degreeCelsius}.

Soil pH and electrical conductivity (EC) were determined for all samples using \SI{15}{\gram} of soil. The soil pH was measured using a WTW pH 340 (WTW GmbH, Weilheim, Germany) with a Sentix 81 electrode. EC was measured using a conductivity meter (LE703, Mettler Toledo, United States). For inorganic ion analyses, \SI{5}{\gram} of each soil sample was dried for \SI{12}{\hour} at \SI{50}{\degreeCelsius} and ground to a finer consistency. Then, the leaching method described by \citet{Genderjahn2018} was employed to extract the ions from the soil. Subsequently, the resulting leachate was measured using ion chromatography (IC). The anions investigated included chloride, nitrate, phosphate, and sulfate, while the cations analyzed were sodium and calcium. Ion concentrations were calculated based on standards (Roth, Multi-Element IC-Standard Solution) using ChromStar7 software. All analyses were performed in triplicate.

\subsection{DNA extraction and quantitative PCR analysis}

TTotal genomic DNA was extracted in duplicates from \SI{0.5}{\gram} of soil for PA and \SI{0.25}{\gram} for SG using the Power Soil DNA extraction kit (Qiagen, Hilden, Germany), following the protocol described by the manufacturer. DNA extraction was exclusively performed on samples subjected to treatments involving microorganisms (in situ, biocrust, and plant treatments), resulting in a total of 116 samples. Sterile treatments devoid of biological relevance were excluded from the DNA extraction. The abundance of bacterial subunit 16S rRNA gene copies was determined by quantitative PCR (qPCR), following the procedure of \citet{Bernhard2018}. qPCR assays were performed on a CFX 96 Connect Real-Time System (Bio-rad Laboratories, CA, United States) in 96-well plates. The primers Eub341F (5\('\)-ATT ACC GCG GCT GCT GG-3\('\)) and Eub534R (5\('\)-CCT ACG GGA GGC AGC AG-3\('\)) \citep{Muyzer1993} were used. The qPCR mixture consisted of \SI{10}{\micro\liter} KAPA SYBR Fast qPCR Kit Master Mix (2x) Universal (Kapa Biosystems, Sigma-Aldrich, Germany), \SI{0.4}{\micro\liter} of each forward and reverse primer, \SI{4.2}{\micro\liter} of DNA-free water, and \SI{5}{\micro\liter} of extracted soil DNA template, resulting in a total volume of \SI{20}{\micro\liter}. The PCR conditions were as follows: initial denaturation at \SI{95}{\degreeCelsius} for \SI{3}{\minute}, followed by 40 cycles of denaturation at \SI{95}{\degreeCelsius} for \SI{3}{\second}, annealing at \SI{60}{\degreeCelsius} for \SI{20}{\second}, and elongation at \SI{72}{\degreeCelsius} for \SI{30}{\second}. We performed a melting curve analysis from \SI{65}{\degreeCelsius} to \SI{95}{\degreeCelsius}, with an increment of \SI{0.5}{\degreeCelsius} to ensure the qPCR specificity. We used standard curves from a known number of PCR-amplified target genes of a pure culture of \textit{Escherichia coli} to estimate the unknown bacterial abundance. All the samples were analyzed in three replicate reactions.

\subsection{Sequencing}

For sequencing, DNA extracted from each sample was amplified using the primer pair 515F (5\('\)-\texttt{GTG YCA GCM GCC GCG GTA A}-3\('\)) and 806R (5\('\)-\texttt{GGA CTA CNV GGG TWT CTA AT}-3\('\)), targeting the V4 hypervariable regions of the bacterial 16S rRNA gene \citep{Apprill2015, Parada2016} including additional barcodes. The polymerase chain reactions were performed in triplicate in a total volume of \SI{25}{\micro\liter}, following the conditions described in \citet{Rodriguez2022} with modifications (reducing the number of denaturation cycles to 20 instead of 30). To confirm the amplification success, PCR products were analyzed by electrophoresis in \SI{1.5}{\percent} agarose gels stained with Gelred\textsuperscript{TM} Nucleic acid gel stain (Biotium, United States). The triplicate PCR samples were pooled and then purified using the carboxyl-coated magnetic beads (Agencourt\textsuperscript{\textregistered} AMPure\textsuperscript{\textregistered} XP Kit, Beckman Coulter, Brea, CA, United States). The purified DNA was quantified using the Qubit Fluorometer (Invitrogen\textsuperscript{TM}, Thermo Fisher Scientific, United States). For sequencing, all PCR products were pooled in equal concentrations of \SI{20}{\nano\gram} for Paired-end sequencing ($2\times 300$\,bp) at Eurofins (Eurofins Genomics Europe Shared Services GmbH, Ebersberg, Germany).

\subsection{Processing of 16S rRNA data}

To generate the amplicon sequence variants (ASVs), the sequencing library was first demultiplexed using cutadapt v3.4 \citep{Martin2011} with the following parameters: \texttt{-e 0.2 -q 15,15 -m 150 --discard-untrimmed}. Only read pairs with the correct barcodes at both ends were retained for further analysis. Next, the DADA2 package v1.20 \citep{Callahan2016} in R v4.1 was employed to generate ASVs from the trimmed reads using a pooled approach with the following parameters: \texttt{truncLen = c(240,200), maxN = 0, rm.phix = TRUE, minLen = 200}. Taxonomic assignment was performed using DADA2 and the SILVA database v138 \citep{Quast2012}. To filter out non-target sequences, ASVs representing chloroplasts, mitochondria, and singletons were removed from the dataset. The raw Illumina sequencing data have been deposited at the European Nucleotide Archive under the BioProject ID \texttt{PRJEB60029}.

\subsection{Data analysis}

All statistical analyses were conducted in R Studio v4.0.5. We performed the Kruskal-Wallis test to determine differences in soil properties, the relative abundance of 16S rRNA, and diversity indices. Post-hoc pairwise comparisons were performed using the Mann–Whitney test, with \(p\)-values adjusted using the Bonferroni correction for multiple comparisons.

To minimize bias due to sequencing depth, all samples were rarefied to 12,000 reads per sample using the \texttt{rtk} package \citep{Saary2017}. Alpha diversity indices, including the Shannon index, observed species, and evenness, were calculated using the \texttt{vegan} package \citep{Oksanen2013}. Relative abundances were calculated by transforming absolute read numbers, and soil physicochemical properties were normalized by subtracting the mean and dividing by the standard deviation. Nonmetric multidimensional scaling (NMDS) with Bray-Curtis distance dissimilarity was performed to assess differences in bacterial structure. The significance of differences between sites, horizon, and time was estimated using PERMANOVA (\texttt{adonis} function) and post-hoc pairwise comparisons (\texttt{pairwise.adonis}) from the \texttt{pairwiseAdonis} package \citep{MartinezArbizu2017} with a significance of \(p < 0.05\). Soil properties influencing bacterial composition were determined using distance-based redundancy analysis (dbRDA) with the \texttt{capscale} function in \texttt{vegan}. The multicollinearity of these variables was examined using the variance inflation factor (VIF), and values \(>10\) were removed. Finally, the FAPROTAX database \citep{Louca2016} was used to analyze the whole bacterial community and predict function from taxonomy.

Correlation analyses between molecular analysis and soil properties were performed using Pearson correlation (\(p < 0.05\)). The results were visualized using the \texttt{ggplot2} package \citep{Wickham2011}.

For further analyses, only abundant ASVs with a proportion of total sequences greater than \SI{0.05}{\percent} were retained to reduce underrepresented ASVs and table complexity. Indicator species were identified using the \texttt{indval} function in the \texttt{labdsv} package to find ASVs leading to changes in multivariate patterns \citep{Roberts2016}. Associations were considered significant with an R-value of 0.8 (\(p < 0.05\)). Then, a network analysis was conducted to understand the distribution patterns of bacterial assemblages using a simplified version of the ecological interactions \citep{Barberan2012, Qian2018}. The co-occurrence network was constructed using the \texttt{igraph} package \citep{Csardi2006} and visualized through the Fruchterman-Reingold layout. Pairwise correlations among ASVs were calculated using the \texttt{rcorr} function within the \texttt{Hmisc} package \citep{Harrell2019}. A Pearson correlation coefficient of R-value \(>0.7\) (\(p < 0.01\)) between two ASVs was considered statistically robust. Topological properties of the network, such as the number of nodes and edges, clustering coefficient, average degree, connectance, and modularity, were calculated using the \texttt{igraph} package. Furthermore, clusters of highly correlated ASVs were grouped into modules using the \texttt{WGCNA} package \citep{Langfelder2008}, and their associations with physicochemical properties were explored. Highly connected nodes (hub ASVs) were then identified from each module.

\section{Results}
\subsection{Soil manipulation experiment}

The microcosms represent a successional gradient of soil development divided into three biological treatments (\cref{fig:figure_1_label}, detailed in Supplementary \cref{fig:supp_figure_s3_label}): soil with \textit{in situ} microorganisms (\textit{in situ}), soil with \textit{in situ} microorganisms and BSCs (BSC), and soil with \textit{in situ} microorganisms and plants (plant). The \textit{in situ} treatment had no plant growth or BSC development, similar to the sterile treatment. In the BSC treatment, varying types and coverage were observed among sites and horizons. Specifically, a leaden BSC developed in PA, while a greenish BSC formed in SG in both horizons. The A horizons of both sites showed nearly complete coverage of the microcosms with BSCs. In contrast, the B horizons showed patchy coverage without forming a compact or extended layer of BSC on the surface. Regarding the plant treatments, similar patterns were observed. In PA, the roots were light brown, and the biomass (both below- and aboveground) and vegetative development were comparatively lower than in SG. In contrast, SG showed dark brown roots, higher biomass, and reached the flowering stage in both horizons. In the PA-Ahor, plants with numerous lateral roots ranged from \SIrange{2}{3}{\centi\metre} in height. Conversely, in PA–Bhor, plants were less than \SI{5}{\milli\metre} tall, had a short taproot with few lateral roots, and experienced a mortality rate of over \SI{50}{\percent} after seven weeks. In SG–Ahor, flowering plants with heights ranging from \SIrange{11}{14}{\centi\metre} were observed. These plants had a \SI{7}{\centi\metre} long taproot and numerous lateral roots covering completely the microcosms. On the other hand, SG–Bhor exhibited plants ranging from \SIrange{1}{3}{\centi\metre} in height, with flower buds and roots reaching \SI{5}{\centi\metre} in length with fewer lateral roots than SG–Ahor. After removing the plants at 12 weeks, no visual differences were observed in the roots between weeks 12 and 16, which remained compact and defined with no signs of degradation. Given the intact nature of the roots during this period, it is considered that these conditions did not impact the short-term degradation of this organic material. Overall, the differences between treatments were less pronounced in the B horizons than in the A horizons, indicating more favorable conditions for BSC and plant development in the A horizons.

\subsection{Soil physical and chemical properties}

Soil samples were analyzed to determine the mass distribution of macro- and microaggregates and the concentrations of C\textsubscript{t} and N\textsubscript{t} in each soil aggregate size (Supplementary \cref{tab:supp_table_s2_label}). No significant differences in aggregate size distribution were found between the study sites (PA--SG; \(n = 54\)), with average proportions of \SI{67}{\percent} macroaggregates, \SI{27}{\percent} large microaggregates, and \SI{7}{\percent} small microaggregates. However, significant differences existed between horizons within each site (Ahor--Bhor; \(n = 27\), \(p < 0.01\)). In both sites, the proportion of macroaggregates was highest in the B horizon, while microaggregates dominated the A horizon. Specifically, macroaggregates mean values were \SI{54}{\percent} in PA--Ahor, \SI{81}{\percent} in PA--Bhor, \SI{63}{\percent} in SG--Ahor, and \SI{70}{\percent} in SG--Bhor. For microaggregates, the means were \SI{23}{\percent} in PA--Ahor, \SI{9}{\percent} in PA--Bhor, \SI{19}{\percent} in SG--Ahor, and \SI{15}{\percent} in SG--Bhor (The full dataset is available in Supplementary \cref{tab:supp_table_s2_label}). Significant differences in the concentration of C\textsubscript{t} and N\textsubscript{t} were observed between sites and horizons (\(p < 0.01\)). C\textsubscript{t} was higher in macro- and large microaggregates of PA (mean \SI{1.2}{\percent}) than in SG (mean \SI{0.7}{\percent}), while N\textsubscript{t} was significantly higher in macro- and microaggregates of SG (mean \SI{0.08}{\percent}) than in PA (mean \SI{0.03}{\percent}). Analysis between horizons showed that macro- and microaggregates of PA--Bhor had a higher C\textsubscript{t} mean of \SI{2}{\percent} than PA--Ahor’s \SI{0.8}{\percent}. Conversely, SG--Ahor’s macro- and large microaggregates had a higher C\textsubscript{t} mean of \SI{1.1}{\percent} than SG--Bhor’s \SI{0.8}{\percent}. In terms of N\textsubscript{t}, higher percentages were observed in SG--Ahor (mean \SI{0.1}{\percent}) and PA--Ahor (mean \SI{0.033}{\percent}) than in SG--Bhor (mean \SI{0.07}{\percent}) and PA--Bhor (mean \SI{0.026}{\percent}).

Over time, we observed changes in aggregate size distribution as well as in C\textsubscript{t} and N\textsubscript{t} concentrations (Supplementary \cref{tab:supp_table_s2_label}). In PA--Ahor, the C\textsubscript{t} of large microaggregates decreased from \SI{0.93}{\percent} in T0 to \SI{0.83}{\percent} in T3. In PA--Bhor, the proportion of macroaggregates increased from \SI{67}{\percent} to \SI{81}{\percent}, while for large microaggregates decreased from \SI{28}{\percent} to \SI{14}{\percent} between T0 and T3. Consequently, MWD increased from \SI{0.8}{\milli\metre} in T0 to \SI{0.94}{\milli\metre} in T3. In SG--Ahor, no marked contrast was observed from T0 to T3, while in SG--Bhor, the percentage of C\textsubscript{t} in macroaggregates increased from \SI{0.2}{\percent} to \SI{0.3}{\percent} over time. Regarding differences between treatments, microorganism-plant treatments showed a noticeable increase in macroaggregates and, thus, in MWD compared to the sterile treatments at T3 (\cref{fig:figure_2_label}). The MWD values for sterile compared to plant treatments were as follows: PA--Ahor (mean \SI{0.7}{mm} vs. \SI{0.78}{\milli\metre}), PA--Bhor (mean \SI{0.91}{mm} vs. \SI{0.95}{\milli\metre}), SG--Ahor (mean \SI{0.75}{mm} vs. \SI{0.87}{\milli\metre}), and SG--Bhor (mean \SI{0.87}{mm} vs. \SI{0.89}{\milli\metre}).

Regarding TOC and N\textsubscript{bulk} (without aggregate size separation) in the bulk soil samples, SG exhibited significantly higher levels (\SI{0.7}{\percent} and \SI{0.06}{\percent}, respectively) than PA (\SI{0.1}{\percent} and \SI{0.02}{\percent}, respectively). When considering the horizons, both N\textsubscript{bulk} and TOC were significantly higher in PA--Ahor (\SI{0.03}{\percent} and \SI{0.14}{\percent}, respectively) and SG--Ahor (\SI{0.08}{\percent} and \SI{0.9}{\percent}, respectively) than PA--Bhor (\SI{0.01}{\percent} and \SI{0.06}{\percent}, respectively) and SG--Bhor (\SI{0.04}{\percent} and \SI{0.46}{\percent}, respectively). Over time, a notable increase in TOC was observed only in SG--Bhor, rising from \SI{0.32}{\percent} in T0 to \SI{0.48}{\percent} in T3. No clear trends were identified for the treatments (\cref{fig:figure_2_label}, Supplementary \cref{tab:supp_table_s2_label}).

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M3-Figure_2.jpg}
	\caption{Physicochemical properties in Pan de Azúcar (PA) and Santa Gracia (SG) show total organic carbon (TOC), nitrogen in bulk soil (Nbulk), pH, electrical conductivity (EC), and mean weight diameter (MWD) in three-time points of the soil manipulation experiment. Time is represented by T0 (original soil sample), T1 (2 weeks), and T3 (16 weeks). Different colors represent the different treatments. Each bar is the mean of biological triplicates.}
	\label{fig:M3-F2}
\end{figure}
